---
title: "Understanding the data"
author: "Emilia Pankowska"
date: "4 listopada 2016"
output: 
  pdf_document: 
    fig_height: 4
    fig_width: 5
---
```{r, echo = FALSE, include = FALSE}
library("dplyr")
library("ggplot2")
library("pander")
panderOptions("table.split.table", 120)
theme_set(theme_bw())

```



# Loading the data

It is assumed that working directory is *PearsonIOKI_DataScientist* frolder. The "data.csv" should be in working directory.

During loading the data empty strings are replaced with NA.
```{r}
df = read.table("../data.csv", sep =",", head = TRUE, na.strings = "")
```

## Getting look at the data

```{r}
head(df)
```

```{r}
str(df)
```

```{r}
summary(df)
```



# Cleaning the data

The result of *summary* function shows that:

1. there are some missing data in *country* and *avg_score* variables
2. Variables *avg_score*, *completion* and *inv_rate* are between 0 and 1 -- so there is no outlier values or typos.

The values of factor variables can be listed to check if there is anything unexpected.

```{r}
apply(df[, 2:4], 2, function(x) sort(table(x)))
```

In the given set, there is no missing values or values that would look suspicious.

## Dealing with missing data

Because there is only few missing cases, it is possible to print them: 

```{r, echo = FALSE}
df_miss <- df[!complete.cases(df),]
rownames(df_miss) = c()
pander(df_miss)
```

There are two learners with missing country Checking if it's possible to get this info from data:
```{r, echo = FALSE}
miss_country <- unique(df[is.na(df$country), 1])
unique(df[df$learner_id %in% miss_country,1:2 ])
```


Because there are only ten (of over 80K) cases with missing values, there is no point of replacing them and they can be ignored.



## Prepering data to analyse

There is a script *clean_data.R* that contains all data manipulation needed to be done, according to preaper raw data from *data.csv* to analyse. The cleaned data.frame is saved in file *clean_data.RData*


```{r, echo = FALSE}
load('../clean_data.RData')
```

# Some plots
```{r, echo = FALSE, fig.height= 3, fig.width=5}
ggplot(df, aes(x = completion)) + 
  geom_histogram(binwidth = 0.05) + 
  labs(list(title = "Completion", y = "no. of cases"))
```

```{r, echo = FALSE, fig.height= 3, fig.width=5}
ggplot(df, aes(x = avg_score)) + 
  geom_histogram(binwidth = 0.05) + 
  labs(list(title = "Average score", y = "no. of cases"))
```

```{r, echo = FALSE, fig.height= 3, fig.width=5}
ggplot(df, aes(x = inv_rate)) + 
  geom_histogram(binwidth = 0.05) + 
  labs(list(title = "Deviation from suggested order", y = "no. of cases"))
```

```{r, echo = FALSE}
label_facet <- c("f" = 'study alone',
                 "t" = 'study in course')

ggplot(df, aes(x = completion)) + 
  geom_histogram(binwidth = 0.05, aes(y=..ncount..)) + 
  facet_grid(in_course~., labeller = as_labeller(label_facet)) +
  labs(list(title = "Completion", y = "normalized no. of cases"))
```

```{r, echo = FALSE}
ggplot(df, aes(x = avg_score)) + 
  geom_histogram(binwidth = 0.05, aes(y=..ncount..)) + 
  facet_grid(in_course~., labeller = as_labeller(label_facet)) +
  labs(list(title = "Average score", y = "normalized no. of cases"))
```

```{r, echo = FALSE}
ggplot(df, aes(x = inv_rate)) + 
  geom_histogram(binwidth = 0.05, aes(y=..ncount..)) + 
  facet_grid(in_course~., labeller = as_labeller(label_facet)) +
  labs(list(title = "Deviation from suggested order", y = "normalized no. of cases"))
```

