---
title: "Understanding the data"
author: "Emilia Pankowska"
date: "4 listopada 2016"
output: pdf_document
---
```{r, echo = FALSE, include = FALSE}
library("dplyr")
library("ggplot2")
theme_set(theme_bw())

```



# Loading the data

The "data.csv" should be in working directory.

Empty strings are replaced with NA 
```{r}
df = read.table("data.csv", sep =",", head = TRUE, na.strings = "")
```

## Getting look at the data

```{r}
head(df)
```

```{r}
str(df)
```

```{r}
summary(df)
```



# Cleaning the data

## Checking *strange* data
The result of *summary* function shows that:

1. there are some missing data in *country* and *avg_score* variables
2. Variables *avg_score*, *completion* and *inv_rate* are between 0 and 1 -- there is no outlier values or typos.

The values of factor variables can be listed to check if there is anything unexpected.

```{r}
apply(df[, 2:4], 2, function(x) sort(table(x)))
```
All values make senes. 
There is a  

## Dealing with missing data

Because there is only few missing cases, it is possible to print them: 

```{r}
df[!complete.cases(df),]
```

There are two learners with missing country Checking if it's possible to get this info from data:
```{r}
miss_country <- unique(df[is.na(df$country), 1])
unique(df[df$learner_id %in% miss_country,1:2 ])
```


Because there are only ten (of over 80K) cases with missing values, there is no point of replacing them and they can be ignored.

Removing missing data:
```{r}
df <- df[complete.cases(df), ]
```

# Some plots
```{r}
ggplot(df, aes(x = completion)) + geom_histogram(binwidth = 0.05)
ggplot(df, aes(x = avg_score)) + geom_histogram(binwidth = 0.05)
ggplot(df, aes(x = inv_rate)) + geom_histogram(binwidth = 0.05)


ggplot(df, aes(x = completion)) + geom_histogram(binwidth = 0.05, aes(y=..ncount..)) + facet_grid(in_course~.)
ggplot(df, aes(x = avg_score)) + geom_histogram(binwidth = 0.05, aes(y=..ncount..)) + facet_grid(in_course~.)
ggplot(df, aes(x = inv_rate)) + geom_histogram(binwidth = 0.05, aes(y=..ncount..)) + facet_grid(in_course~.)
```

