---
title: "Students"
author: "Emilia Pankowska"
date: "6 listopada 2016"
output: 
  pdf_document: 
    fig_height: 3.5
    fig_width: 5
---

The data in this file were aggregated to learner_id. The *avg_score*, *avg_completion* and *avg_inv_rate* are the mean value of *avg_score*, *completion* and *inv_rate* respectively.

```{r, echo = FALSE, include = FALSE}
library("dplyr")
library("ggplot2")
theme_set(theme_bw())


load("../clean_data.RData")

df_l <- df %>%
  group_by(learner_id, country, in_course) %>%
  summarize(no_units = n(),
            no_reviews = sum(as.numeric(as.vector(is_review))),
           #dist_no_units = n_distinct(unit),
            avg_score = mean(avg_score),
            avg_completion = mean(completion),
            avg_inv_rate = mean(inv_rate)) %>%
  ungroup()
```


## Country of origin

The countries with less than 60 learners have been gathered together in 'other' group.

Each bar is labeled by percentage of all students.

```{r, echo = FALSE}
cc <- sort(table(df_l$country)[table(df_l$country) >= 60], decreasing = TRUE)

t = geom_text(aes(y = ((..count..)/sum(..count..)),
                  label = scales::percent((..count..)/sum(..count..))), 
              stat = "count", 
              vjust = -0.25,
              size = 2.5,
              color = 'orange')

df_l %>% 
  mutate(country_bar = ordered( (ifelse(country %in% rownames(cc), as.character(country), 'other')),c(rownames(cc), 'other'))
         # , country_order = ifelse(country %in% rownames(cc), cc[as.character(country)], 0 )
         )%>%
  # arrange(desc(country_order)) %>%       
  ggplot(aes(x = country_bar)) + 
  geom_bar() + 
  labs(list(title = "Number of learners in countries", x = "country", y = "no. of learners")) + t
```

## Number of units taken by students
Each bar is labeled by percentage of all students.

```{r, echo = FALSE}

  ggplot(df_l, aes(x = as.factor(no_units))) + geom_bar()+labs(list(title = "", x = "no. of units", y = "no. of learners")) +  t
  

```

## Average score and the number of units taken by student
Blue line is a linear treand.

```{r, echo = FALSE }
ggplot(df_l, aes(as.factor(no_units),avg_score)) + 
  geom_boxplot() + 
  geom_smooth(method = "lm", se=FALSE, color="blue", aes(group=1))+labs(list(title = "", x = "no. of units", y = "average score"))
```


