---
title: "Analysis of online workbook"
author: "Emilia Pankowska"
output: 
  beamer_presentation: 
    colortheme: orchid
    fig_height: 5
---

## Introduction

```{r, echo = FALSE, include = FALSE}
library("dplyr")
library("ggplot2")
library("pander")
library("tidyr")
panderOptions("table.split.table", 70)
theme_set(theme_bw())

load("../clean_data.RData")
```

The dataset comes from a Pearson e-learning platform, an online workbook that is used alongside a paper textbook. The dataset contains a following information:

* learner_id 
* country
* unit
* in_course -- "t" if the student belongs to course taught by a teacher; "f" if study alone
* avg_score -- average percentage score on all activities within a given unit
* copletion -- the percentage of activities completed in a given unit, out of all activities available in that unit
* inv_rate -- this is the extent to which a student deviates from the suggested order of activities by the pedagogy experts within a given unit. A value of zero indicates no departure from the suggested order, a value of one indicates a complete reversal of the order.



## Introdaction 

  

```{r, echo = FALSE}
no_students = length(unique(df$learner_id))
no_countries = length(unique(df$country))
table1 <- df %>% 
  group_by(in_course) %>%
  summarize(no_of_students = n_distinct(learner_id),
            percentage = paste(round(no_of_students*100.0 / no_students ,2), "%", sep = ""))%>%
  mutate(course = ifelse(in_course == 't', 'course with teacher', 'study alone')) %>%
  arrange(desc(in_course)) %>% 
  select (course, no_of_students, percentage)
  
```

The dataset contains information about `r format(no_students, big.mark = " ")` students from `r format(no_countries, big.mark = " ")` counties.
\tiny
```{r, echo = FALSE}
pander(format(table1, big.mark = " "))
```



## Number of units taken by students

```{r, echo = FALSE}
df_l <- df %>%
  group_by(learner_id, country, in_course) %>%
  summarize(no_units = n(),
            no_reviews = sum(as.numeric(as.vector(is_review))),
           #dist_no_units = n_distinct(unit),
            avg_score = mean(avg_score),
            avg_completion = mean(completion),
            avg_inv_rate = mean(inv_rate)) %>%
  ungroup()

t = geom_text(aes(y = ((..count..)/sum(..count..)),
                  label = scales::percent((..count..)/sum(..count..))), 
              stat = "count", 
              vjust = -0.25,
              size = 4,
              color = 'white')

label_facet <- c("f" = 'study alone',
                 "t" = 'course with teacher')

ggplot(df_l, aes(x = as.factor(no_units))) + 
  geom_bar() +
  labs(list(title = "", x = "no. of units", y = "no. of learners")) +
  t

```

## Number of units taken by students

```{r, echo=FALSE}

ggplot(df_l, aes(x = no_units)) + 
  geom_histogram(binwidth = 0.5, aes(y = ..ncount..)) + 
  facet_grid(in_course~., labeller = as_labeller(label_facet))+
  labs(list(title = "", x = "no. of units", y = "no. of learners"))
```

\tiny
```{r, echo=FALSE}

df_l %>% mutate(u = cut(no_units,c(0,3,6,9,12,17), c("1-3","4-6", "7-9","10-12", "13-17") ))%>%
  group_by(in_course, u) %>% 
  summarize(no_of_students = n()) %>%
  mutate( percent = 
          round(ifelse(in_course == 'f', no_of_students*100.0/ table1[[2,2]], no_of_students*100.0/ table1[[1,2]] ),2)) %>%
  select(in_course, u, percent) %>%
  spread (u, percent) %>% pander()
        
```

## Number of units taken by students
It can be seen that only small percantage of self-study students took more then first three units (63% to 37%), when in other group most of the students took more than three units (36% to 65%).

## Completion, average score and inv_rate  

There is small (but significant) difference in mean values of Completion, average score and inv_rate within two groups: *study alone* and *course with teacher*. 

Between two groups, there is small but significant difference between mean values of Completion, average score and inv_rate.
The 'study alone' group has lower *avg_score* and *completion* but higher *inv_rate* than a 'course with teacher' group.


## Completion
```{r, echo = FALSE, fig.height=3, fig.width = 5}
label_facet <- c("f" = 'study alone',
                 "t" = 'course with teacher')

ggplot(df, aes(x = completion)) + 
  geom_histogram(binwidth = 0.05, aes(y=..ncount..)) + 
  facet_grid(in_course~., labeller = as_labeller(label_facet)) +
  labs(list(title = "Completion", y = "normalized no. of cases"))
```

\tiny
```{r, echo = FALSE}
comp <- data.frame(do.call("rbind", by(df$completion, df$in_course, function(x) round(summary(x),2 ))))
colnames(comp) <- c("Min." ,"1st.Qu.","Median", "Mean","3rd.Qu.", "Max." )
rownames(comp) <- c('study alone','course with teacher' )
comp = format(comp, round = 2)
pander(comp)
```

## Average score

```{r, echo = FALSE, fig.height=3, fig.width = 5}
ggplot(df, aes(x = avg_score)) + 
  geom_histogram(binwidth = 0.05, aes(y=..ncount..)) + 
  facet_grid(in_course~., labeller = as_labeller(label_facet)) +
  labs(list(title = "Average score", y = "normalized no. of cases"))
```
\tiny
```{r, echo = FALSE}
comp <- data.frame(do.call("rbind", by(df$avg_score, df$in_course, function(x) round(summary(x),2 ))))
colnames(comp) <- c("Min." ,"1st.Qu.","Median", "Mean","3rd.Qu.", "Max." )
rownames(comp) <- c('study alone','course with teacher' )
pander(comp)
```

## Inv_rate
```{r, echo = FALSE, fig.height=3, fig.width = 5}
ggplot(df, aes(x = inv_rate)) + 
  geom_histogram(binwidth = 0.05, aes(y=..ncount..)) + 
  facet_grid(in_course~., labeller = as_labeller(label_facet)) +
  labs(list(title = "Deviation from suggested order", y = "normalized no. of cases"))
```
\tiny
```{r, echo = FALSE}
comp <- data.frame(do.call("rbind", by(df$inv_rate, df$in_course, function(x) round(summary(x),2 ))))
colnames(comp) <- c("Min." ,"1st.Qu.","Median", "Mean","3rd.Qu.", "Max." )
rownames(comp) <- c('study alone','course with teacher' )
pander(comp)
```

## Analysis per unit

Between units there is no big difference in *avg_score* and *inv_rate*, but the *completion* varies: between units and reviews and it decreases with further units.

## Analysis per unit


```{r, echo=FALSE, fig.height=3.5  }
ggplot(df, aes(unit,completion)) + 
  geom_boxplot() + 
  geom_smooth(color="blue", se=FALSE, aes(group=1)) + 
  geom_smooth(method = "lm", se=FALSE, color="green", aes(group=1)) +
  facet_grid(in_course~., labeller = as_labeller(label_facet))
```

```{r, echo=FALSE, fig.height=3.5 }
ggplot(df, aes(unit,avg_score)) + 
  geom_boxplot() + 
  geom_smooth(color="blue", se=FALSE, aes(group=1)) + 
  geom_smooth(method = "lm", se=FALSE, color="green", aes(group=1)) +
 facet_grid(in_course~., labeller = as_labeller(label_facet))
```

## Analysis per unit

```{r, echo=FALSE, fig.height=3.5  }
ggplot(df, aes(unit,inv_rate)) + 
  geom_boxplot() + 
  geom_smooth(color="blue", se=FALSE, aes(group=1)) + 
  geom_smooth(method = "lm", se=FALSE, color="green", aes(group=1)) +
  facet_grid(in_course~., labeller = as_labeller(label_facet))

```
